---
title: "Text Analysis"
author: "Joi Chu-Ketterer"
date: "May 28, 2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

```{r, message=FALSE, warning=FALSE}
library(pdftools)
library(tm)
library(stringr) #let's me extract the substrings
library("qdapRegex") # let's me extract quotations
library("dplyr") #need for arrange and for inner join
library("tidytext") #for ngrams and lexicons
library("tidyverse") #for separate()
```

Reading in the PDF file and checking page count
```{r}
text <- pdf_text('tlm.pdf')
pdf_info("tlm.pdf")$pages
```

Cleaning out punctuation
```{r}
text2 <- gsub('\n',' ',text)
text3 <- gsub('[[:punct:] ]+',' ',text2)
text3 <- tolower(text3)
```

1 ----- Punctuation Frequency

Extracting punctuation and saving as dataframe
```{r}
punct <- gsub('[0-9]+', ' ', text2)
punct <- tolower(punct)
punct <- gsub("[a, b, c, d, e, f, g, h, i, j, k, l, m, n, o, p, q, r, s, t, u, v, w, x, y, z, -]", " ", punct)
punct <- strsplit(punct, " ")

punct_table <- table(unlist(punct))

punct_df <- as.data.frame(punct_table)
punct_df <- punct_df[punct_df$Var1!="",]
```

Cleaning dataframe and distributing counts correctly
```{r}
punct_df_clean <- punct_df %>%
    mutate(Freq = case_when(
      Var1 == '—"' ~ Freq + 0,
      Var1 == ';' ~ Freq + 2,
      Var1 == ';"' ~ Freq + 0,
      Var1 == ':' ~ Freq + 0,
      Var1 == '!' ~ Freq + 1,
      Var1 == '!"' ~ Freq + 0,
      Var1 == '?' ~ Freq + 6,
      Var1 == '?"' ~ Freq + 0,
      Var1 == '.' ~ Freq + 21,
      Var1 == '."' ~ Freq + 0,
      Var1 == "'" ~ Freq + 0,
      Var1 == '"' ~ Freq + 40
    ))
```

Removing unecessary rows and writing as a csv file
```{r}
punct_df_clean = punct_df_clean[-c(1,3,6,8,10),]
names(punct_df_clean) <- c("punctuation", "count")

punct_df_clean$percent <- (punct_df_clean$count/sum(punct_df_clean$count))*100

write.csv(punct_df_clean, "punctuation_percent.csv", row.names = FALSE)
```

2 ----- QUOTE RATIO

Extracting quotes
```{r}
#using text since that still has the punctuation 
quotes <- vector(mode = "list", length = 1)

quote <- rm_between(text, '"', '"', extract=TRUE)

for (i in quote)
{ quotes <- c(quotes, i)
  
}

#final list of all quotes
quotes = unlist(quotes)
```

Saving as dataframe
```{r}
quote_df <- data.frame(quotes = unlist(quotes))
quote_df_clean <- dplyr::filter(quote_df,  !is.na(quotes))
```

Cleaning dataframe
```{r}
for (i in quote_df_clean){
  i <- gsub('[[:punct:] ]+',' ', i)
}

quote_df_clean$quotes <- sapply(quote_df_clean$quotes, function(x) gsub('[[:punct:] ]+',' ', x))

quote_df_clean$quotes <- sapply(quote_df_clean$quotes, function(x) gsub('\n',' ', x))
```

Calculates word count per quote, and finds sum
```{r}
quote_df_clean$count <- sapply(quote_df_clean$quotes, function(x) length(unlist(strsplit(as.character(x), " "))))

sum(quote_df_clean$count)
```

number of words in a quote: 1849
number of words in total: 9251 (discovered later)

Saving quote ratio dataframe
```{r}
type <- c("Quote", "Total")
count <- c(1849, 9251)
quote_ratio <- data.frame(type, count)

quote_ratio$percentage <- (quote_ratio$count/sum(quote_ratio$count))*100

write.csv(quote_ratio, "quote_ratio.csv", row.names = FALSE)
```

3 ----- SENTENCE LENGTH

Extracting sentences and saving as a dataframe
```{r}
sentence <- unlist(strsplit(text, '\\.'))
sentence_df <- data.frame(sentence)
sentence_df_clean <- dplyr::filter(sentence_df,  !is.na(sentence))
```

Cleaning the sentence dataframe
```{r}
for (i in sentence_df_clean){
  i <- gsub('[[:punct:] ]+',' ', i)
}

sentence_df_clean$sentence <- sapply(sentence_df_clean$sentence, function(x) gsub('[[:punct:] ]+',' ', x))

sentence_df_clean$sentence <- sapply(sentence_df_clean$sentence, function(x) gsub('\n',' ', x))

sentence_df_clean$sentence <- gsub('[0-9]+', '', sentence_df_clean$sentence)

sentence_df_clean[1,1] <-sapply(sentence_df_clean[1,1], function(x) gsub('THE LITTLE MERMAID','', x))
```

Calculates word count
```{r}
sentence_df_clean$count <- sapply(sentence_df_clean$sentence, function(x) length(unlist(strsplit(as.character(x), " "))))
sentence_df_clean <- sentence_df_clean[sentence_df_clean$sentence!=" ",]
```

Saving as CSV for external visualization tools
```{r}
write.csv(sentence_df_clean, "sentence_df_clean.csv", row.names = FALSE)
```

4 ----- WORD FREQUENCY

Extracts individual words and creates table
```{r}
words <- strsplit(text3, " ")

freq <- table(unlist(words))
```

Saves table as a dataframe and calcultes sum
```{r}
count <- cbind(names(freq),as.numeric(freq))
word_count <- as.data.frame(count)
names(word_count) <- c("Word", 'Count')
word_count$Word <- levels(word_count$Word)[word_count$Word]
word_count$Word <- gsub('[0-9]+', '', word_count$Word)
word_count <- word_count[word_count$Word!="",]
   
word_count$Count <- levels(word_count$Count)[word_count$Count]
word_count$Count <- as.numeric(word_count$Count)

sum(word_count$Count)
```

Removing stop words for dynamic visualization
```{r}
nostop <- removeWords(text3, stopwords("en"))
```

Cleans and calculates frequency
```{r}
words_nostop <- strsplit(nostop, " ") 
table_nostop <- table(unlist(words_nostop))

count_nostop <- cbind(names(table_nostop),as.numeric(table_nostop))
count_nostop <- as.data.frame(count_nostop)
names(count_nostop) <- c("Word", 'Count')
count_nostop$Word <- levels(count_nostop$Word)[count_nostop$Word]
count_nostop$Word <- gsub('[0-9]+', '', count_nostop$Word)
count_nostop <- count_nostop[count_nostop$Word!="",]
count_nostop <- count_nostop[count_nostop$Word!="s",]
   
count_nostop$Count <-levels(count_nostop$Count)[count_nostop$Count]
count_nostop$Count <- as.numeric(count_nostop$Count)
```

Displays word counts in descending order
```{r}
count_nostop_sorted <- arrange(count_nostop, desc(Count))
```
  
Graph 1 - recreated in Tableau
```{r}
pdf("words_top_ten.pdf")

top_n(count_nostop_sorted, n=10, Count) %>%
        ggplot(aes(x = reorder(Word, Count), Count)) +
        geom_bar(stat="identity", fill="lightseagreen") + coord_flip() + scale_y_continuous(expand = c(0,0), limits = c(0,70)) + labs(title = "Word Count", subtitle = expression(paste("top ten words used in ", italic("The Little Mermaid"))), x = " ", y = "") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "white")) + theme(axis.ticks = element_blank())

dev.off()
```

Saves top ten words as CSV for external visualization
```{r}
top_ten_df <- top_n(count_nostop_sorted, n=10, Count)

write.csv(top_ten_df, "top_ten_df.csv", row.names = FALSE)
```

Graph 2 - recreated in Tableau
```{r}
pdf( "words_more_than_ten.pdf")

subset(count_nostop_sorted, Count>10)  %>%
ggplot(aes(x = reorder(Word, Count), Count)) + geom_bar(stat="identity", fill="lightseagreen") + coord_flip() + scale_y_continuous(expand = c(0,0), limits = c(0,70)) + labs(title = "Word Count", subtitle = expression(paste("words used more than ten times in ", italic("The Little Mermaid"))), x = " ", y = "") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_rect(fill = "transparent",colour = NA),plot.background = element_rect(fill = "transparent",colour = NA), axis.line = element_line(colour = "white")) + theme(axis.ticks = element_blank())

dev.off()
```

Saves words used more than ten times as CSV for external visualization
```{r}
more_ten_df <- subset(count_nostop_sorted, Count>10)

write.csv(more_ten_df, "more_ten_df.csv", row.names = FALSE)
```

5 ----- SENTIMENT ANALYSIS

Finds sentiment for each word

```{r}
bing <- get_sentiments("bing")

sentiment <- inner_join(count_nostop_sorted, bing, by = c('Word' = 'word'))

names(sentiment) <- c("Text", 'Count', "Sentiment")
sentiment <- arrange(sentiment, desc(Count))
```

Calculating sentiment weight
```{r}
sentiment_weight <- sentiment %>%
  count(Text, Sentiment, wt = Count) %>%
  spread(Sentiment, n, fill = 0) %>%
  mutate(Sentiment = positive - negative) %>%
  arrange(Sentiment)

sentiment_weight <- sentiment_weight %>%
  count(Sentiment, Text, wt = Sentiment)
```

Graph 3 - sentiment contribution for words used more than five times
```{r}
sentiment_weight %>%
  count(Sentiment, Text, wt = Sentiment) %>%
  filter(abs(n) >= 5) %>%
  mutate(Text = reorder(Text, n)) %>%
  ggplot(aes(Text, Sentiment, fill = n)) +
  geom_bar(stat = "identity") +
    labs(x = NULL, y = NULL, title = "Sentiment Contribution", 
         subtitle = expression(paste("by word counts greater than five in ", italic("The Little Mermaid")))) +
  scale_fill_gradient2(low='#6276bc', mid='#cbe9e2', high='#98d3c6', space='Lab', name = "Contribution")+ theme_classic() + theme(axis.text.x = element_text(angle = 90, hjust = 1), panel.background = element_rect(fill = "transparent",colour = NA),plot.background = element_rect(fill = "transparent",colour = NA), legend.key = element_blank())

ggsave("sentiment_contribution.png", bg = "transparent")
```

Calculating overall sentiment count and relative percentage and saving as dataframe
```{r}
overall_sentiment <- data.frame(table(sentiment$Sentiment))
names(overall_sentiment) <- c("Sentiment", 'Count')
overall_sentiment$Percent <- (overall_sentiment$Count / sum(overall_sentiment$Count)) *100
```

Calculating attributes for donut graph
```{r}
overall_sentiment$ymax <- cumsum(overall_sentiment$Percent)
overall_sentiment$ymin <- c(0, head(overall_sentiment$ymax, n=-1))
overall_sentiment$labelPosition <- (overall_sentiment$ymax + overall_sentiment$ymin) / 2

overall_sentiment$label <- paste(overall_sentiment$Sentiment, "\n",overall_sentiment$Percent,'%')
```

Graph 4 - break down of negative and positive sentiment by word
```{r}
pdf("testing.pdf")

mycolors <- c("#623d4f", "#4f623d")
            
ggplot(overall_sentiment, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=Sentiment)) +
  geom_rect() +
  geom_text( x=2, aes(y=labelPosition, label=label, color=Sentiment), size=6) + 
  coord_polar(theta="y") +
  scale_fill_manual(values = mycolors) +
  scale_color_manual(values = mycolors) +
  xlim(c(-1, 4)) +
  theme_void() +
  theme(legend.position = "none") + labs(title = "Sentiment Breakdown", subtitle = expression(paste(italic("The Little Mermaid "), "is generally a positive story")))

dev.off()
```

6 ----- GENDER ANALYSIS

Saving text3 as a dataframe 
```{r}
text_df <- as.data.frame(text3)
text_df$text3 <- gsub('[0-9]+', '', text_df$text3)
names(text_df) <- c("Text")
```

Tokenizing words into bigrams 
```{r}
tokenized <- text_df %>% 
  unnest_tokens(word, Text, token = "ngrams", n=2)
```

Filtering bigrams by leading word that start with "she" or "he"
```{r}
pronouns <- c("he", "she")

bigram_counts <- tokenized %>%
  count(word, sort = TRUE) %>%
  separate(word, c("word1", "word2"), sep = " ") %>%
  filter(word1 %in% pronouns) %>%
  #n is the count, how many times it appears in the text
  count(word1, word2, wt = n, sort = TRUE)
```

Calculating leading gender counts
```{r}
leading_counts <- as.data.frame(sum(bigram_counts$n, bigram_counts$word1 == "she"))
names(leading_counts) <- c("she")
leading_counts$he <- sum(bigram_counts$n, bigram_counts$word1 == "he")
```

Calculating leading gender ratios
```{r}
ratios <- bigram_counts %>%
    group_by(word2) %>%
    filter(sum(n) > 3) %>%
    ungroup() %>%
    spread(word1, n, fill = 0) %>%
    #calculates the ratio, adding 1 since some of the counts are 0
    mutate_if(is.numeric, list(~(. + 1) / sum(. + 1))) %>%
    mutate(logratio = log2(she / he)) %>%
    arrange(desc(logratio))  
```

Graph 5 - ratio frequency of words following she and he
```{r}
ratios %>%
    mutate(abslogratio = abs(logratio)) %>%
    group_by(logratio < 0) %>%
    top_n(15, abslogratio) %>%
    ungroup() %>%
    mutate(word = reorder(word2, logratio)) %>%
    ggplot(aes(word, logratio, color = logratio < 0)) +
    geom_segment(aes(x = word, xend = word,
                     y = 0, yend = logratio), 
                 size = 2, alpha = 0.6) +
    geom_point(size = 0) +
    coord_flip() +
    labs(x = NULL, y = NULL, title = "Bigram Frequency", subtitle = expression(paste("Trailing words paired with 'he' and 'she' in ", italic("The Little Mermaid")))) +
    scale_color_discrete(name = "", labels = c("More 'she'", "More 'he'")) +
    scale_y_continuous(breaks = seq(-3, 3),
                       labels = c("0.125x", "0.25x", "0.5x", 
                                  "Same", "2x", "4x", "8x")) + theme_classic() + theme(panel.background = element_rect(fill = "transparent",colour = NA),plot.background = element_rect(fill = "transparent",colour = NA), legend.key = element_blank()) 

ggsave("gender_analysis.png", bg = "transparent")
```




```{r}